Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-07-13 writeup/rocha_etal_supporting-info.tex
Small fix
Dimtris' comments

Table \ref{tab:llc} shows details of the LLC spin-up hierarchy.
Both simulations analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting in 2011, and the 16 most significant  tidal components.
The LLC control files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}).

@@

Figure \ref{figA1} shows KE frequency spectra for the LLC simulations and mooring data.
At the uppermost common depth (20 m), very low frequencies (periods $>$ 10 days)
and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
data are more energetic at intermediate frequencies. At 40 m, the LLC simulations and observations
have consistent spectra at periods larger than about 6 h. The data from the moorings
are more energetic at higher frequencies,
except for the first two harmonics that are apparent in the KESS data at 40 m.
There are at least three reasons for the slightly more energetic super-inertial
variability in the observations: 1) The high-
frequency variability is dominated by high-frequency noise due to mooring vertical
excursions; 2) The model does not have enough resolution to resolve the full internal
wave spectrum; 3) The model is not forced at frequencies higher than 6h.

@@

 \begin{figure}
    \begin{center}
      \setfigurenum{S1}
      \includegraphics[width=40pc]{figs/fig_s2_3.pdf}
   \caption{A comparison between spectra of the two LLC simulations 1/24$^\circ$
   (green) and 1/48$^\circ$ (purple). The statistical errorbars represent conservative
   95$\%$ significance levels.  As expected the
   higher-resolution simulation (LLC4320) essentially extends the spectra of the
   lower-resolution simulation (LLC2160) towards smaller scale, demonstrating that
   the higher resolution is resolving a wider array of scales. Between 20 and 100 km
   the spectra of the two simulations are visually indistinguishable.}
   \label{figS2_3}
   \end{center}
 \end{figure}

@@

 \begin{figure}
    \begin{center}
      \setfigurenum{S4}
      \includegraphics[width=40pc]{figs/fig_s2_1.pdf}
   \caption{Potential density: a comparison between Roemmich-Gilson Argo
           climatology [updates from \cite{roemmich_gilson2009}] (a, b) and the LLC simulations 1/24$^\circ$ (c, d)
           and  1/48$^\circ$ (e, f). Of course, the Argo climatology is smoother
           owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce well the strong
           seasonality in the stratification. The upper ocean is weakly stratified in late winter/early fall
           and re-stratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
           at all seasons (not shown). The Argo climatology is available online at
           \texttt{http://sio-argo.ucsd.edu/RG\_Climatology.html}.}
   \label{figS2_1}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S5}
      \includegraphics[width=40pc]{figs/figA1.pdf}
   \caption{A comparison between frequency spectra of LLC simulations and two
   available moored current meter records.}
   \label{figA1}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S6}
      \includegraphics[width=40pc]{figs/fig_s3.pdf}
   \caption{The seasonal variability of the WOA 2013 \citep{levitus_etal2013} stratification averaged  over
   the domain and the associated three gravest pressure modes. Only the upper 450 m is shown.
   Clearly, the changes to the stratification
   are dramatic, with the formation of a seasonal pycnocline in summer. The strong
   seasonality of upper-ocean stratification yields a strong seasonality in the
   near-surface shape and amplitude of baroclinic pressure modes.
   In particular, the surface amplitude is much large in summer. Thus the projection
   of baroclinic tides on the surface may vary significantly seasonally. The WOA data is
   available online at \texttt{https://www.nodc.noaa.gov/OC5/woa13/}.}
   \label{figS3}
   \end{center}
 \end{figure}

 \begin{table}
 \label{tab:llc}
 \settablenum{S1}
 \caption{\small The MITgcm Latitute-Longitude Cap (LLC) spin-up hierarchy. This study analyzes
          the output of LLC2160 and LLC4320. Model fields are available hourly.
          Adapted from \cite{rocha_etal2016}.}
 \begin{center}
 \begin{tabular}{ | c | c | c | c | c |}
 \hline
 Simulation & Resolution & Time-step & Period  & Tides \\ \hline
 ECCO2 Adjoint & $1/6^{^\circ}$ & 1200 s & January 2009 - December 2011 & No\\
 LLC 1080   & $1/12^{^\circ}$ & 90 s & January 2010 - July 2012 & No\\
 LLC 2160   &  $1/24^{^\circ}$ & 45 s  & January 2011 - April 2013 & Yes\\
 LLC 4320   &  $1/48^{^\circ}$ & 25 s & September 2011 - September 2012 & Yes \\
 \hline
 \end{tabular}
 \end{center}
 \end{table}

>>>>

Table \ref{tab:llc} shows details of the LLC spin-up hierarchy.
Both simulations analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting in 2011, and the 16 most significant  tidal components.
The LLC code and model configuration files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}).

@@

Figure \ref{figA1} shows KE frequency spectra for the LLC simulations and mooring data.
At the uppermost common depth (20 m), very low frequencies (periods $>$ 10 days)
and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
data are more energetic at intermediate frequencies. At 40 m, the LLC simulations and observations
have consistent spectra at periods larger than about 6 h. The data from the moorings
are more energetic at higher frequencies,
except for the first two harmonics that are apparent in the KESS data at 40 m.
There are at least two reasons for the slightly more energetic super-inertial
variability in the observations: 1) The high-
frequency variability is dominated by high-frequency noise due to mooring vertical
excursions; and 2) The model does not have enough resolution to resolve the full internal
wave spectrum.

@@

 \begin{figure}
    \begin{center}
      \setfigurenum{S1}
      \includegraphics[width=40pc]{figs/fig_s2_3.pdf}
   \caption{Comparison between spectra of the 1/24$^\circ$
   (green) and 1/48$^\circ$ (purple) LLC simulations. The statistical error bars represent conservative
   95$\%$ significance levels.  As expected the
   higher-resolution simulation (LLC4320) essentially extends the spectra of the
   lower-resolution simulation (LLC2160) towards smaller scales, demonstrating that
   the higher resolution is resolving a wider array of scales. Between 20 and 100 km
   the spectra of the two simulations are visually indistinguishable.}
   \label{figS2_3}
   \end{center}
 \end{figure}

@@

 \begin{figure}
    \begin{center}
      \setfigurenum{S4}
      \includegraphics[width=40pc]{figs/fig_s2_1.pdf}
   \caption{Potential density: a comparison between Roemmich-Gilson Argo
           climatology [updates from \cite{roemmich_gilson2009}] (a, b) and the 1/24$^\circ$ (c, d)
           and  1/48$^\circ$ (e, f) LLC simulations. Of course, the Argo climatology is smoother,
           owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce the strong
           seasonality in the stratification. The upper ocean is weakly stratified in late winter/early fall
           and re-stratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
            during all seasons (not shown). The Argo climatology is available online at
           \texttt{http://sio-argo.ucsd.edu/RG\_Climatology.html}.}
   \label{figS2_1}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S5}
      \includegraphics[width=40pc]{figs/figA1.pdf}
   \caption{Comparison between frequency spectra of LLC simulations and two
   available moored current meter records.}
   \label{figA1}
   \end{center}
 \end{figure}

 \begin{figure}
    \begin{center}
      \setfigurenum{S6}
      \includegraphics[width=40pc]{figs/fig_s3.pdf}
   \caption{The seasonal variability of the WOA 2013 \citep{levitus_etal2013} stratification averaged  over
   the domain and the associated three gravest pressure modes. Only the upper 450 m is shown.
   Clearly, the changes to the stratification
   are dramatic, with the formation of a seasonal pycnocline in summer. The strong
   seasonality of upper-ocean stratification yields a strong seasonality in the
   near-surface shape and amplitude of baroclinic pressure modes.
   In particular, the surface amplitude is much larger in summer. Thus the projection
   of baroclinic tides on the surface may have significant seasonal variations. The WOA data is
   available online at \texttt{https://www.nodc.noaa.gov/OC5/woa13/}.}
   \label{figS3}
   \end{center}
 \end{figure}

 \begin{table}
 \label{tab:llc}
 \settablenum{S1}
 \caption{\small The MITgcm Latitute-Longitude Cap (LLC) spin-up hierarchy. This study analyzes
          the output of LLC2160 and LLC4320. Model fields are available hourly.
          Adapted from \cite{rocha_etal2016}.}
 \begin{center}
 \begin{tabular}{ | c | c | c | c | c |}
 \hline
 Simulation & Grid spacing & Time-step & Period  & Tides \\ \hline
 ECCO2 Adjoint & 18 km & 1200 s & January 2009 - December 2011 & No\\
 LLC 1080   & $1/12^{^\circ}$ & 90 s & January 2010 - June 2012 & Yes \\
 LLC 2160   &  $1/24^{^\circ}$ & 45 s  & January 2011 - April 2013 & Yes\\
 LLC 4320   &  $1/48^{^\circ}$ & 25 s & September 2011 - October 2012 & Yes \\
 \hline
 \end{tabular}
 \end{center}
 \end{table}
